Robertson and Cao Project

### 2024-03-16 code/analysis-code/analysis1.qmd
Added analysis2.qmd, added more to analysis1

---
title: "Data Analysis and Modelling of Teen Vaccination Surveys"
date: "`r format(Sys.Date()'"
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence
---
First, I will start by setting an equation for the generalized logistic regression model with all of the predictor variables intended for use.
```{r}
levels(mydata$P_UTDHPV)
mydata2 <- droplevels(mydata[!mydata$P_UTDHPV == 'Missing Data', ]) # I will do some brief cleaning and remove all rows with missing values for our response variable, 'P_UTDHPV', to perform a binomial regression
str(mydata2$P_UTDHPV) # Check for removal of missing values
fitall <- glm(P_UTDHPV ~ AGE + SEX + STATE + INS_STAT2_I + INCQ298A + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + RENT_OWN + FACILITY, data = mydata2, family = binomial) # define the equation for the binomial logistic regression
```

@@

Now I will start the step-wise comparison in both the forwards and backwards direction to allow the model to add or drop predictors depending on the responding AIC value.
```{r}
step(fitstart, direction = "both", scope = list(lower = ~1, upper = ~ AGE + SEX + STATE + INS_STAT2_I + INCQ298A + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + RENT_OWN + FACILITY), data = mydata, family = binomial) # define the lower and upper limits of the step wise comparison 
```
The model chosen includes the predictors STATE, FACILITY, RACEETHK, AGE, EDUC1, INS_BREAK_I, INCPOV1, SEX, LANGUAGE, MOBIL_1, and INS_STAT2_I. 
Because of the extremely high residual value, I will perform some diagnostics to determine what is the issue. First, I will look at a residual plot to check for non linearity. Next, I will use a VIF analysis to check for multicollinearity. I am suspicious of co linearity as many of these socioeconomic predicts might be inherently related.

>>>>

---
title: "Data Analysis and Modelling of Teen Vaccination Surveys"
date: '`r format(Sys.Date())`'
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence
---
First, I will start by setting an equation for the generalized logistic regression model with all of the predictor variables intended for use.

The model aims to predict the vaccination status (P_UTDHPV) based on various predictor variables like age, sex, state, insurance status, income, race/ethnicity, education, language, etc. The response variable P_UTDHPV represents whether a teenager is up-to-date on HPV vaccination.
```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","cleandata1.rds")

#load data
mydata <- readRDS(data_location)

levels(mydata$P_UTDHPV)
mydata2 <- droplevels(mydata[!mydata$P_UTDHPV == 'Missing Data', ]) # I will do some brief cleaning and remove all rows with missing values for our response variable, 'P_UTDHPV', to perform a binomial regression
str(mydata2$P_UTDHPV) # Check for removal of missing values
fitall <- glm(P_UTDHPV ~ AGE + SEX + STATE + INS_STAT2_I + INCQ298A + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + RENT_OWN + FACILITY, data = mydata2, family = binomial) # define the equation for the binomial logistic regression
```

> A Stepwise model comparison is performed to iteratively add or remove predictor variables based on their significance in improving the model fit.

@@

Now I will start the step-wise comparison in both the forwards and backwards direction to allow the model to add or drop predictors depending on the responding AIC value. This is performed with the `step()` function, setting up the scope for the variables to consider.
```{r}
step(fitstart, direction = "both", scope = list(lower = ~1, upper = ~ AGE + SEX + STATE + INS_STAT2_I + INCQ298A + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + RENT_OWN + FACILITY), data = mydata, family = binomial) # define the lower and upper limits of the step wise comparison 
```
The model chosen includes the predictors STATE, FACILITY, RACEETHK, AGE, EDUC1, INS_BREAK_I, INCPOV1, SEX, LANGUAGE, MOBIL_1, and INS_STAT2_I. 
Because of the extremely high residual value, I will perform some diagnostics to determine what is the issue. First, I will look at a residual plot to check for non linearity. Next, I will use a VIF analysis to check for multicollinearity. I am suspicious of co linearity as many of these socioeconomic predicts might be inherently related.

### 2024-03-16 products/manuscript/Manuscript.qmd
Added analysis2.qmd, added more to analysis1

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

@@

We start our exploratory analysis by displaying the distribution of our predictors, including the socioeconomic variables that we chose for analysis.
The distribution of the numerical income to poverty ratio can be shown below.
It displays a value from 1-3.
A ratio displaying less than 1 depicts an income less than the poverty level.Any income ratio \>1 indicates an income greater than the poverty level.
Most of the observations were gathered from households with income ratios of 3, as shown by the skewed plot below.
![Figure 1. Income-poverty ratio distribution](results/figures/income.png)

@@

![Figure 8. Vaccine status distribution](results/figures/State-vaccine-completion.png)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(here)
```

@@

We start our exploratory analysis by displaying the distribution of our predictors, including the socioeconomic variables that we chose for analysis.
The distribution of the numerical income to poverty ratio can be shown below.
It displays a value from 1-3.
A ratio displaying less than 1 depicts an income less than the poverty level.Any income ratio \>1 indicates an income greater than the poverty level.
Most of the observations were gathered from households with income ratios of 3, as shown by the skewed plot below.

![Figure 1. Income-poverty ratio distribution](/results/figures/income.png)

@@

![Figure 8. Vaccine status distribution](../results/figures/State-vaccine-completion.png)
